---
title: "Quiz2"
author: "Kien Pham"
date: "2024-02-02"
output: html_document
---
#QUIZ 2 MAT115

##1 
```{r}
sum <- 0
for (x in 1:99) {
  sum = sum + 1 + .01*x
}
1 + sum + 2
```

##2
Load the dataset olive from dslabs package. 

a. How many olive oils are from the Coast-Sardinia area with the percentage of stearic acid less than 2.35? 

b. Find the second highest percentage of stearic acid for the olive oils from the Coast-Sardinia area. 

c. Among all olive oils from the Coast-Sardinia area, find the percentage of eicosenoic acid for the olive oil with the highest percentage of stearic acid.
```{r}
library(dslabs)
library(tidyverse)

str(olive)

olive |> 
  filter(area == "Coast-Sardinia" & stearic < 2.35) 

olive[which(olive$area == "Coast-Sardinia" & olive$stearic < 2.35),]

olive_CSd <- olive[which(olive$area == "Coast-Sardinia"),]
rank_stearic <- rank(-olive_CSd$stearic)
olive_CSd[which(rank_stearic == 2),]

olive_CSd$eicosenoic[which(rank_stearic == 1)]
```


